---
title: "AE 03: Duke Forest + data visualization"
---

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(openintro)
```

## Exercise 1

Suppose you're helping some family friends who are looking to buy a house in Duke Forest.
As they browse Zillow listings, they realize some houses have garages and others don't, and they wonder: **Does having a garage make a difference?**

Luckily, you can help them answer this question with data visualization!

-   Make histograms of the prices of houses in Duke Forest based on whether they have a garage.
    -   In order to do this, you will first need to create a new variable called `garage` (with levels `"Garage"` and `"No garage"`).
    -   Below is the code for creating this new variable. Here, we `mutate()` the `duke_forest` data frame to add a new variable called `garage` which takes the value `"Garage"` if the text string `"Garage"` is detected in the `parking` variable and takes the test string `"No garage"` if not.

```{r}
#| label: garage
#| eval: true

duke_forest <- duke_forest |>
  mutate(garage = if_else(str_detect(parking, "Garage"),   "Garage", "No garage"))
```

-   Then, facet by `garage` and use different colors for the two facets.
-   Choose an appropriate binwidth and decide whether a legend is needed, and turn it off if not.
-   Include informative title and axis labels.
-   Finally, include a brief (2-3 sentence) narrative comparing the distributions of prices of Duke Forest houses that do and don't have garages. Your narrative should touch on whether having a garage "makes a difference" in terms of the price of the house.

```{r}
#| label: price-garage

ggplot(duke_forest, aes(x = price, fill = garage)) +
  geom_histogram(binwidth = 50000) +
  facet_wrap(~ garage, ncol = 1) +
  labs(
    title = "Distribution of House Prices in Duke Forest by Garage Presence",
    x = "Price (USD)",
    y = "Count"
  ) +
  theme(legend.position = "none")
```

The distribution of house prices for homes with a garage appears to be shifted slightly to the right compared to those without a garage, indicating that homes with garages generally have higher prices. Both distributions are right-skewed, but the "Garage" category has a longer tail extending into the higher price ranges. Thus, having a garage does seem to be associated with a higher house price.

::: callout-important
Now is a good time to render, commit, and push.
Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding.
:::

## Exercise 2

It's expected that within any given marker larger houses will be priced higher.
It's also expected that the age of the house will have an effect on the price.
However in some markets new houses might be more expensive while in others new construction might mean "no character" and hence be less expensive.
So your family friends ask: "In Duke Forest, do houses that are bigger and more expensive tend to be newer ones than those that are smaller and cheaper?"

Once again, data visualization skills to the rescue!

-   Create a scatter plot to exploring the relationship between `price` and `area`, conditioning for `year_built`.
-   Use `geom_smooth()` with the argument `se = FALSE` to add a smooth curve fit to the data and color the points by `year_built`.
-   Include informative title, axis, and legend labels.
-   Discuss each of the following claims (1-2 sentences per claim). Your discussion should touch on specific things you observe in your plot as evidence for or against the claims.
    -   Claim 1: Larger houses are priced higher.
    -   Claim 2: Newer houses are priced higher.
    -   Claim 3: Bigger and more expensive houses tend to be newer ones than smaller and cheaper ones.

```{r}
#| label: price-area
#| warning: false

ggplot(duke_forest, aes(x = area, y = price, color = year_built)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "black") + # Using black for the trend line to distinguish from colored points
  labs(
    title = "Price vs. Area of Houses in Duke Forest",
    x = "Area (sq ft)",
    y = "Price (USD)",
    color = "Year Built"
  )
```

-   **Claim 1: Larger houses are priced higher.**
    The scatter plot shows a strong positive correlation between area and price. As the area on the x-axis increases, the price on the y-axis tends to increase as well, supporting the claim.

-   **Claim 2: Newer houses are priced higher.**
    Looking at the color gradient, lighter blue points (indicating more recent years) appear throughout the price range, but there isn't a strict separation where *all* newer houses are higher. However, there is a visible trend where many of the higher-priced homes (top right) are lighter in color (newer).

-   **Claim 3: Bigger and more expensive houses tend to be newer ones than smaller and cheaper ones.**
    The cluster of points in the upper-right corner (high price, large area) consists largely of lighter blue points, indicating they are newer. The lower-left corner (low price, small area) contains a mix but has many darker blue points (older homes). This supports the claim that the biggest and most expensive homes tend to be newer construction.

::: callout-important
Now is a good time to render, commit, and push.
Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding.
:::
